ms_5

### 2020-03-20 ms_5.Rmd
ms 5 complete!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(janitor)
library(tidyverse)
```

*Copy pasted from last week
```{r}

@@

*New for this week
```{r}
# For this week, I am going to plot the Drug Overdose Deaths in April of 2015 by
# state according to the CDC. I want to use this data just because it is the
# first recorded date of drug overdose deaths. I am hoping there will be some
# expected trend across states.

@@

@@

ggplot(drugodapr15, aes(x= reorder(state_name,-data_value), y= data_value)) +
geom_col() +
labs(title = "Drug Overdose Deaths by State (April 2015)", 
     x = "State", 
     y = "Drug Overdose Deaths") +
theme_classic()+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}

# I am going to do the same with July 2019, which is the latest date of
# recording.

@@

```{r}
difference_over_time<- drugodapr15 %>%
  full_join(drugodjul19, by = "state_name") %>%
  mutate(difference = data_value.y - data_value.x)

ggplot(difference_over_time, aes(x= reorder(state_name,-difference), y= difference)) +
geom_col() +
labs(title = "Difference in Drug Overdose Deaths between April 2015 and July 2019", 
     x = "State", 
     y = "Difference in Drug Overdose Deaths") +
theme_classic()+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# I want to load all necessary packages.

library(dplyr)
library(janitor)
library(tidyverse)
```

*Reading in drug overdose data was copy pasted from last week
```{r, echo = FALSE}

@@

*The graphs below are new for this week. I don't know that I am satisfied with the readability of the last graph and will probably change it slightly in my final version.

```{r, echo = FALSE}

# For this week, I am going to plot the Drug Overdose Deaths in April of 2015 by
# state according to the CDC. I want to use this data just because it is the
# first recorded date of drug overdose deaths. I am hoping there will be some
# expected trend across states.

# I am not sure if this makes a difference but I am going to change the variable
# to be numeric (from factor) to rank later on.

@@

# Here, I want to rank the deaths by state in descending order and filter them
# such that we are only looking at info from April of 2015.

@@

# Now, I want to ggplot the data. In the aesthetics, when I simply plot the
# state name vs data value, the data values do not rank in descending order
# despite specifying earlier, so I used the trick from PSet 3 of reordering in
# the x specifications.

ggplot(drugodapr15, aes(x= reorder(state_name,-data_value), y= data_value)) +
geom_col() +

# I want to label the title and x and y axes so that they are more descriptive.

labs(title = "Drug Overdose Deaths by State (April 2015)", 
     x = "State", 
     y = "Drug Overdose Deaths") +

# I changed the theme such that the backing of the graph is blank.

theme_classic()+

# The x axis details are cluttered so I chose to angle them for easy reading.

theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}

# I am going to do the same with July 2019, which is the latest date of
# recording. All code is the same except that the filter specifications at the
# start are for July of 2019 and I changed the title to specify July of 2019.

@@

```{r, echo = FALSE}

# I now want to look at how drug overdose death levels changed between 2019 and
# 2015 by state. To do so, I combined the two data tables from above using a
# full join.

difference_over_time<- drugodapr15 %>%
  full_join(drugodjul19, by = "state_name") %>%

# I also want to add a column which indicates the difference in deaths between
# the two data tables.

mutate(difference = data_value.y - data_value.x)

# I want to graph the difference over time by state with the x axis ranking the
# difference decrease.

ggplot(difference_over_time, aes(x= reorder(state_name,-difference), 
                                 y= difference)) +
geom_col() +

# I also changed the title label and the y axis label to specify the
# differences.

labs(title = 
    "Difference in Drug Overdose Deaths between April 2015 and July 2019", 
     x = "State", 
     y = "Difference in Drug Overdose Deaths") +

# I want the cleanest theme.

theme_classic()+

# I changed the labelling on the x axis to be less cluttered.

theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
